### 2017-03-27 ThesisText/background.tex
Adds parts for background

E-Learning is usually provided over the Internet using platforms
called~\gls{lms}. Examples of~\gls{lms} include Moodle, Blackboard Learn, D2L,
among others. These platforms organize users by separating them into roles such
as Professors and Students or Authors and Learners. The Professors or Authors
are the users who organize courses in universities or training in companies,
create and publish content, perform evaluation of students, etc. Students or
Learners are the users which actually perform the training and follow courses
in the platforms.~\cite{ind_010, ind_011, ind_013}

@@

% =============================================================================
\section{Data in Learning Management Systems}
\TODO{This section should just be part of the previous one.}

Resources and activities, as well as the logs detailing their usage, are also
kept in a database. The usage of these systems by professors and students
generates entries.

@@

Data Mining aims at finding patterns in data by looking at the different
features of the data entries. These patterns allows then to make non trivial
assumptions in data.~\cite{book_dm_practical}

To find these patterns we build and train model such as Decision Trees,
Bayesian Networks, etc, from the data.

Data is usually stored in databases which may follow different paradigms and
have different architectures. But generally, we refer to data entries as
datasets. We also refer to a group of datasets as data repository.

\TODO{Talk about specific models}

>>>>

E-Learning is usually provided over the Internet using platforms
called~\gls{lms}. Examples of~\gls{lms} include Moodle, Blackboard Learn, D2L,
among others. These platforms organize users by separating them into roles such
as Professors and Students or Authors and Learners. The Professors or Authors
are the users who organize courses in universities or training in companies,
create and publish content, perform evaluation of students, etc. Students or
Learners are the users which actually perform the training and follow courses
in the platforms.

Resources and activities, as well as the logs detailing their usage, are also
kept in a database. The usage of these systems by professors and students
generates entries.~\cite{ind_010, ind_011, ind_013}

@@

@@

Data Mining aims at finding patterns in data by looking at the different
features of the data entries. These patterns allows us to make non trivial
assumptions in data.~\cite{book_dm_practical}

Data is usually stored in databases which may follow different paradigms and
have different architectures. But generally, we refer to data entries as
datasets. We also refer to a group of datasets as a data repository.

To find patterns in data, models are trained either using supervised or
unsupervised learning. The following sections describe these concepts and
introduce models from both.

\subsection{Supervised Learning}

Supervised learning is done when data made up of a collection of labeled
objects which have a list of features and a label, usually called class. The
idea is that a model is trained using instances from labeled data in order to
correctly classify unlabeled instances later.

There exist several supervised learning models, but in this project we use only
two. They are Decision Trees and Naive Bayes
Classifiers.~\cite{book_ml,book_dm_practical}

\subsubsection{Decision Trees}

Decision trees are a model that make use of a tree to classify an object. The
model works by sorting an object down the tree, taking a single feature into
account on each node of the tree. The branches on each node are the possible
values of the feature being taken into account on that node. The leafs of the
tree assign a class to the object.

Decision trees are generally best suited for situations where the data has
discrete values. This is the case in this project, since many features are
either integer numbers with a limited range of values, or strings and
enumerated types which are nominal in nature.

There are a few algorithms used to train these model. They are ID3, C4.5 which
is an optimization of ID3, and CART.~\cite{book_ml,book_dm_practical}

\subsubsection{Naive Bayes Classifiers}

Naive Bayes Classifiers are probabilistic models which are based on Bayesian
Networks. A Bayesian Network is a graph that describes directional the
dependencies between features of objects. Those features are described as
random variables.

The models that make up a Naive Bayes Classifier are a particular case of
Bayesian Networks. In these models there is a single node of the network which
is the class of the object. Every feature depends directly on the class and
only on the class.

Each model essentially describes a conditional probability distribution like:

\begin{equation} \label{eq:prob} P(\; x_1, \; \dots, \; x_n \; | \; C),
\end{equation}

where $ C $ is the class of the object and $ x_1, \; \dots, \; x_n $ are the
features of it. The model is able to classify an object by directly applying
the Bayes Theorem to the distribution, obtaining a value a $ C $.

There are different types of Naive Bayes Classifier. Each type is adequate for
different types of features. For example, Guassian Naive Bayes Classifier
describe relations between variables, like equation~\ref{eq:prob}, as normal
distributions. Bernoulli Naive Bayes Classifier are used if the features are
boolean.

Multinomial Naive Bayes Classifier are used when the data has discrete values
and is nominal in nature. This model is used in this project because the data
it deals with is discrete and nominal.~\cite{book_ml,book_dm_practical}

\subsection{Unsupervised Learning}

Unsupervised learning deals with data that is unlabeled. One approach to
unsupervised learning is clustering. The objective of clustering is to find
groups of unlabeled objects. A group would ideally have objects which features
are similar to objects within that group, and dissimilar to objects in other
groups.

One difficulty of clustering is to find the number of clusters that optimally
divide the data. Some algorithms take a number of clusters as an arguments,
while others are able to find an optimal value.

By it's own nature, it may not be possible to find clusters in data, that is,
there may not be any great division of objects given their features in a given
dataset. After executing clustering algorithms a performance measure is used to
determine if that division is good or not.

In this project, the performance measure used is called Silhouette Score.
Silhouette score yields a value between -1 a 1. The interpretation given to
this value is that higher values mean that objects are correctly placed in
their clusters, lower values mean the objects are in the wrong clusters. Values
near zero indicate that the clusters overlap.~\cite{book_ml,book_dm_practical}

\subsubsection{K-Means}

K-Means is an clustering algorithm that find $ K $ clusters in data. The $ K $
variable is given to the algorithm as an argument, hence the name K-Means. The
algorithm works by creating $ K $ random points based on the features of the
data. The algorithm then iterates updating the location of those points until
they converge.~\cite{book_ml,book_dm_practical}

\subsubsection{Affinity Propagation}

Affinity Propagation is another clustering algorithm, but this one does not
take a number of clusters as an argument. This algorithm finds the optimal
number of clusters.~\cite{book_ml,book_dm_practical}

\section{Cross Validation}

The model used in this project and described in this section always go through
two processes. Training and testing. To do this two sets of data are needed,
one for each process. Data for the two processes is usually taken from the same
dataset in different proportions. For example, some projects may take 75\% or
90\% of the data to do training, and use the remainder to test.

Testing a model with different objects then the ones used in training is very
important. This is because a model, while training, will learn the very
specific details of that data and may over fit. When over fitted, a model will
not work for general examples of the data for which it was trained and will
only ``remember'' training data instead.

A way of dividing a dataset into training and testing sets is to do Cross
Validation. Cross Validation divides the data into $ N $ folds. From the
original $ N $ folds, a single one is taken for testing and the model is
trained with the remaining $ N - 1 $ folds.

A model is trained and tested $ N $ times, each time with a different fold
taken for testing. Like this we get $ N $ different executions for the same
dataset and same model and we are able to compare the best, worst, and average
scenario in terms of training and testing.~\cite{book_dm_practical}

### 2017-03-27 ThesisText/experiments.tex
Adds parts for background

Some fields on dataset~\ref{tab:dat_001} are strings and enumerated types,
therefore, these are nominal fields. Because of this characteristic, Decision
Trees and Bayesian Networks are used.~\TODO{Citation is needed on claim that
Trees and BNets are good for nominal data}

@@

    \caption
        [Clustering number of approved courses based on Moodle usage results]
        {Clustering number of approved courses based on Moodle usage results. The table displays the results for the silhouette score for clustering.}

@@

\TODO{These values are the same for every variant of the dataset. This needs to be explained}

>>>>

Some fields on dataset~\ref{tab:dat_001} are strings and enumerated types,
therefore, these are nominal fields. Because of this characteristic, Decision
Trees and Bayesian Networks are used.

@@

    \caption
        [Clustering number of approved courses based on Moodle usage results]
        {Clustering number of approved courses based on Moodle usage results.
        The table displays the results for the silhouette score for clustering.
        The values for every variant of the dataset are approximately the same,
        so only one table is shown.}

@@

Table~\ref{tab:exp_004_res} shows the results of this experiment. The
silhouette score shown is approximately the same for each variant of the
dataset, so only one table is shown for every variant.

It is observable that for two clusters, the score is around 0.8, and for three
clusters the score is around 0.75. This shows how these objects are able to be
separated in groups of two or three without having those groups overlapping.

For a number of clusters equal to or greater than four, we start getting scores
bellow 0.7.

The results shown were obtain with the K-Means algorithm. Affinity Propagation
was also executed, but the results obtained were not good, having scores within
0.0 and 0.5.
